Towards a miniKanren with fair search strategies

### 2019-05-27 Paper/paper.tex
save on-going work

miniKanren is a family of relational programming languages.
\citet{Friedman:2005:RS:1121583,friedman_reasoned_2018} introduce
miniKanren and its implementation in \emph{The Reasoned Schemer} 
and \emph{The Reasoned Schemer, 2nd Ed} (TRS2). Although originally written in 
Scheme, miniKanren has been implemented in many other languages, including
multiple ones using the same language. (See miniKanren.org.) This has come
about because of \citet{Hemann_2016}'s work on microKanren, a core of 
miniKanren comprised of only 54 LOC. Also, miniKanren
has been ported to Ocaml, and named OCanren\citep{kosarev2018typed}). 
\citet{byrd2017unified} have 
demonstrated that miniKanren problems that at first reading seem impossible
are indeed possible by using a Scheme interpreter written in miniKanren, thereby
having access to logic variables within Scheme.
miniKanren.org contains the seeds of many interesting problems and 
their solutions.

\NOTE[DPF]{I thought we agreed to use BFSser throughout?  
In that sense it had emerged a few days ago, and now it has
re-emerged.  It is what we are going to use instead of ``current''.}
\NOTE[LKC]{}

@@

\subsection{fair \texttt{disj}}

@@

\subsection{fair \texttt{conj}}
\label{sec:fairconj}

@@

A more interesting situation is when the first conjunct produces an unbounded
number of states. Consider the following example, a naive specification of 
fair \conj{} 
might require search strategies to produce all sorts of singleton lists, but 
there
would not be any lists of length two or longer, which makes the strategies 
incomplete. 
A search strategy is \emph{complete} if and only if ``every correct answer 
would be discovered after some finite time'' \cite{seres1999algebra}, 
otherwise, it is \emph{incomplete}. In the 
context of miniKanren, a search strategy is complete means that every correct 
answer has a position in large enough answer lists.

@@

Fair DFS (\DFSf) has fair \disj{} and unfair \conj{}. The 
implementation of \DFSf{} differs from \DFSi{}'s in 
\disjtwo{} (Fig.~\ref{fDFS}). The new \disjtwo{} calls a new and 
fair version of \appendInf{}. \appendInfFair{} 
immediately 
calls 
its helper, \texttt{loop}, with the first argument, \texttt{s?}, initialized to 
\texttt{\#{}t}, which indicates that we haven't swapped
\sInf{} and \tInf{}. The swapping 
happens at 
the third \texttt{cond} clause in \texttt{loop}, where \texttt{s?} is updated 
accordingly. The first two \texttt{cond} clauses essentially copy the 
\texttt{car}s and stop recursion when one of the input spaces is obviously 
finite. The third clause, as we mentioned above, is just for swapping. When the 
fourth and last clause runs, we know that both \sInf{} and 
\tInf{} are ended with a thunk, and that we have swapped them. In 
this case, we construct a new thunk. The new thunk swaps back the two spaces in 
the
recursive call to \texttt{loop}. This is unnecessary for fairness---we do it to 
produce answers in a more readable order.

@@

\subsection{improved BFS}

@@

The new type of \texttt{Space} is a pair whose \texttt{car} is a list of 
answers (the bucket), and whose \texttt{cdr} is either \texttt{\#{}f} or a 
thunk returning a space. A falsy \texttt{cdr} means the space is obviously 
finite.

@@

\subsection{compare \BFSimp{} with \BFSser{}}

@@

In general, only \DFSi{} and \DFSbi{} constantly perform well. \DFSf{}
is just as efficient in \appendo, \reverso, and \texttt{quine}s. All BFS
implementations have obvious overhead in most cases. 
Among the three variants of DFS, which all have unfair \conj{}, \DFSf{} is most 
resistant to clause permutation in \texttt{quines}, followed by \DFSbi{} then 
\DFSi{}. \DFSf{}, however, runs out of memory in \texttt{'(I love you)}s. In 
contrast, \DFSbi{} is still more stable than \DFSi{}. Thus, we consider 
\DFSbi{} a competitive alternative to \DFSi{}.
Among the two implementations of \BFS, \BFSimp{} constantly performs as well 
or better. 
It is hard to conclude how \disj{} fairness affects performance based on these 
statistics.
Fair \conj{}, however, imposes considerable overhead constantly except in 
\appendo. The reason might be that those strategies tend to keep more 
intermediate answers in the memory.

@@

Our quantitative evaluation shows that \DFSbi{} is a competitive 
alternative to \DFSi{}, the current miniKanren search strategy,
and that \BFSimp{} is more practical than \BFSser{}.

@@

We acknowledge that we are making large claims about efficiency given that we 
have just a few benchmarks, but our intuition is based on sound principles. ...

\NOTE[DPF]{Finish this paragraph with a clarification and you may
place this anywhere in the conclusion. This may require some consultation with 
Weixi.}
\NOTE[LKC]{I have sent an email to Weixi on the clarification.}

>>>>

\NOTE[LKC]{I have restructured the following paragraph.}

miniKanren is a family of relational programming languages.
\citet{Friedman:2005:RS:1121583,friedman_reasoned_2018} introduce
miniKanren and its implementation in \emph{The Reasoned Schemer} 
and \emph{The Reasoned Schemer, 2nd Ed} (TRS2). After \citet{Hemann_2016} 
describe microKanren, a minimal core of miniKanren comprised of only 54 LOC, 
miniKanren has been implemented in many other languages, including
multiple ones using the same language. (e.g. OCanren\citep{kosarev2018typed}). 
\citet{byrd2017unified} have demonstrated that miniKanren problems that at 
first reading seem impossible are indeed possible by using a Scheme interpreter 
written in miniKanren, thereby having access to logic variables within Scheme.
miniKanren.org contains links to many miniKanren implementations and the seeds 
of many interesting problems and their solutions.

%miniKanren is a family of relational programming languages.
%\citet{Friedman:2005:RS:1121583,friedman_reasoned_2018} introduce
%miniKanren and its implementation in \emph{The Reasoned Schemer} 
%and \emph{The Reasoned Schemer, 2nd Ed} (TRS2). Although originally written in 
%Scheme, miniKanren has been implemented in many other languages, including
%multiple ones using the same language. (See miniKanren.org.) This has come
%about because of \citet{Hemann_2016}'s work on microKanren, a core of 
%miniKanren comprised of only 54 LOC. Also, miniKanren
%has been ported to Ocaml, and named OCanren\citep{kosarev2018typed}). 
%\citet{byrd2017unified} have 
%demonstrated that miniKanren problems that at first reading seem impossible
%are indeed possible by using a Scheme interpreter written in miniKanren, 
%thereby
%having access to logic variables within Scheme.
%miniKanren.org contains the seeds of many interesting problems and 
%their solutions.

\NOTE[DPF]{I thought we agreed to use BFSser throughout?  
In that sense it had emerged a few days ago, and now it has
re-emerged.  It is what we are going to use instead of ``current''.}
\NOTE[LKC]{I am confused. Do you mean we are going to use
	``BFS'' for the search strategy,
	``BFSser'' for Seres's BFS implementation, and
	``the improved BFS'' for our BFS implementation
?}

@@

\subsection{Fair \texttt{disj}}

@@

\subsection{Fair \texttt{conj}}
\label{sec:fairconj}

@@

A more interesting situation is when the first conjunct produces an unbounded
number of states. Consider the following example: a naive specification of 
fair \conj{} 
might require search strategies to produce all sorts of singleton lists, but 
there
would not be any lists of length two or longer, which makes the strategies 
incomplete. 
A search strategy is \emph{complete} if and only if ``every correct answer 
would be discovered after some finite time'' \cite{seres1999algebra}, 
otherwise, it is \emph{incomplete}. In the 
context of miniKanren, a search strategy is complete means that every correct 
answer has a position in large enough answer lists.

@@

Fair DFS (\DFSf) has fair \disj{} and unfair \conj{}. The 
implementation of \DFSf{} differs from \DFSi{}'s in 
\disjtwo{} (Fig.~\ref{fDFS}). The new \disjtwo{} calls a new and 
fair version of \appendInf{}. \appendInfFair{} 
immediately 
calls 
its helper, \texttt{loop}, with the first argument, \texttt{s?}, initialized to 
\texttt{\#{}t}, which indicates that we haven't swapped
\sInf{} and \tInf{}. The swapping 
happens at 
the third \texttt{cond} clause in \texttt{loop}, where \texttt{s?} is updated 
accordingly. The first two \texttt{cond} clauses essentially copy the 
\texttt{car}s and stop recursion when one of the input spaces is obviously 
finite. The third clause, as we mentioned above, is just for swapping. When the 
fourth and last clause runs, we know that both \sInf{} and 
\tInf{} end with thunks, and that we have swapped them. In 
this case, we construct a new thunk. The new thunk swaps back the two spaces in 
the
recursive call to \texttt{loop}. This is unnecessary for fairness---we do it to 
produce answers in a more readable order.

\NOTE[LKC]{Following Boskin's suggestion, ``are ended with a thunk'' -> ``end 
with thunks''}

@@

\subsection{Improved BFS}

@@

The new type of \texttt{Space} is a pair whose \texttt{car} is a list of 
answers (the bucket), and whose \texttt{cdr} is either \texttt{\#{}f} or a 
thunk returning a space. The \texttt{\#{}f} here means the space is obviously 
finite, just like empty list in other implementations.

\NOTE[LKC]{I have rewritten the sentence about \texttt{\#{}f}.}

@@

\subsection{Compare \BFSimp{} with \BFSser{}}

@@

\NOTE[LKC]{(following Boskin's suggestion) constantly -> consistently}

In general, only \DFSi{} and \DFSbi{} consistently perform well. \DFSf{}
is just as efficient in \appendo, \reverso, and \texttt{quine}s. All BFS
implementations have obvious overhead in most cases. 
Among the three variants of DFS, which all have unfair \conj{}, \DFSf{} is most 
resistant to clause permutation in \texttt{quines}, followed by \DFSbi{} then 
\DFSi{}. \DFSf{}, however, runs out of memory in \texttt{'(I love you)}s. In 
contrast, \DFSbi{} is still more stable than \DFSi{}. Thus, we consider 
\DFSbi{} a competitive alternative to \DFSi{}.
Among the two implementations of \BFS, \BFSimp{} constantly performs as well 
or better. 
It is hard to conclude how \disj{} fairness affects performance based on these 
statistics.
Fair \conj{}, however, imposes considerable overhead constantly except in 
\appendo. The reason might be that those strategies tend to keep more 
intermediate answers in the memory.

@@

Our quantitative evaluation shows that \DFSbi{} is a competitive 
alternative to \DFSi{}, the current miniKanren search strategy,
and that \BFSimp{} is more practical than \BFSser{}.
We acknowledge that we are making large claims about efficiency given that we 
have just a few benchmarks, but our benchmarks cover some practically 
interesting cases (e.g. synthesizing expressions that evaluate to \texttt{'(I 
love you)}.

\NOTE[DPF]{Finish this paragraph with a clarification and you may
place this anywhere in the conclusion. This may require some consultation with 
Weixi.}
\NOTE[LKC]{Correcting.}

@@
